A hierarchical approach for modelling X-ray beamlines. Application to a coherent beamline.

### 2019-06-21 paper-hierarchical.tex
some fixes

Accurate calculation and quantitative evaluation of the parameters related to X-ray optics are of great importance for designing, building and exploiting new beamlines. Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential. These calculations allow testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively. The selection of the optimal optical layout requires a detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optics. Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray-tracing techniques. They model the X-ray beam as a collection of incoherent rays. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources can introduce the necessity to complement the ray-tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent hard X-ray beam is still a dream for the new generation storage rings. Despite a considerable increase in the coherent fraction of the new beams, the coherent fraction at typically hard X-ray photon energies is still a few percent for the new sources. Consequently, the ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics. Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually computationally expensive.

@@

 In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented. The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations. Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction. Accuracy can be improved using geometrical methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline. The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding. In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented. The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools. Finally, Section \ref{summary} summarizes the results.

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF-1 and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values are taken from \cite{ESRF2014}, updated for EBS with
    the lattice labelled S28D. 
    %For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. 
    %\todo{Remove ESRF-AT and EBS-S28A} 
    }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\ \\
         %\input{tableElectron.txt}
         ESRF-1  & 387.80  & 3.50  & 10.30 & 1.20 \\ 
         %ESRF-AT  & 414.97  & 3.43  & 10.31 & 1.16 \\ 
         %EBS-S28A  & 27.20  & 3.40  & 5.20 & 1.40 \\ 
         EBS     & 30.18  & 3.64  & 4.37 & 1.37 \\ 
    \end{tabular}
    }
\end{table}

@@

The source size and divergence are usually approximated supposing a Gaussian distribution of the electron positions and divergences (Table \ref{tab:eBeam}) and single-electron photon source size and divergence. In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later. Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the standard deviations of electron beam size and divergence at the center of symmetry of the straight section in a plane perpendicular to the main propagation direction. The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.

@@

Using the values for U18.3, $n$=1, and  $E$=17225~eV (Table \ref{tab:Undulators}) and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca. 8.
%\inred{8 $\approx$ 403.8/50=8.076}
Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5~nm$^2$.

@@

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of  390 $\times$ 900 (H $\times$ V) $\mu$m$^2$. 
We get N.A. values of 2.7 $\times$ 4.9~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 
4.1\% (H) and 40\% (V)for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
0.21\% for ESRF-1 and 2.85\% for EBS, therefore 
$4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}

@@

In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5~nm$^2$ for ESRF-1, 10$\times$5~nm$^2$ for the EBS) the flux ($3.8\cdot10^{12}$ photons/s for ESRF-1 and $8.5\cdot10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction-limited focal spot of  7.75$\times$7.75~nm$^2$. Thus far we ignored the effect of real optics exhibiting surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. It also helps in providing an initial benchmark estimate which should, to a large extent, provide a check of the validity of the numerical calculations that will provide more precise values, but should remain at the same order of magnitude.

@@

We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV and approximated using Gaussian distributions (Eqs.~\ref{eq:photon big sigmas}), considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements. The KB reflector shape are elliptical as required for a point-to-point focusing. For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors. 
The results in Fig.~\ref{fig:ray-tracing}a show two phenomena. One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF-1. Secondly, one can observe a much larger luminosity in the EBS case. The transmission values are 0.16\% for ESRF-1 and 2.22\% for EBS. These values confirm the analytical estimations (0.2\% for ESRF-1 and 2.95\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected. It is therefore expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.

The hybrid method \cite{hybrid} uses geometric ray-tracing combined with wavefront propagation. The code computes diffraction effects when the beam is cropped and can also simulate the effect of mirror figure errors when diffraction is present. The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far-field propagation in all intermediate propagations. If this is not the case, a near field version is available for investigating imaging by individual optics. The interest of the hybrid method is to assess in a fast way the influence of beam coherence because the code is considerably faster than the full methods for dealing with the partial coherence of the synchrotron beam that are described below. Fig.~\ref{fig:ray-tracing}b presents the results using the hybrid model for the beamline with ideal elliptical optical surfaces. When compared with the pure ray-tracing in Fig.~\ref{fig:ray-tracing}a it can be appreciated that the spot is broadened. This is more evident in the vertical direction, because of the smaller focal size. The spot size is about 7-8~nm in V for both storage ring lattices, and in horizontal is about 15~nm for ESRF-1 High$\beta$ and 11~nm for EBS. It is interesting to note the almost round spot that will be obtained with the new EBS lattice.

\begin{figure}
\label{fig:ray-tracing}
\centering
~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
\newline
\includegraphics[width=0.95\textwidth]{GRAPHICS/figure3.png}
\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}. 
a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the BK mirrors) without slope errors. 
b) Images calculated using the hybrid method (ray-tracing plus coherent diffraction) with ideal reflectors with no slope errors. It shows the focal spot degradation due to the diffraction arising from the cropping of the beam by KB mirrors. c) Simulations using the hybrid method including slope errors from the real mirror profiles measured at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}). It can be observed further beam broadening and the growth of satellite structures. 
The focal beam dimensions (FWHM of the histograms) are written in the plot title.
}
\end{figure}

@@

A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that reproduce well the characteristics of the real elements. Once the source is defined, an approach of how best to simulate the beamline optical elements. 
Once the source is defined, several approaches of increasing complexity are used: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}) b) use grazing incidence focusing optics models \cite{Canestrari2014}; and c) introduce the effect of imperfect optics on the simulations, i.e. slope and figure errors to both ideal lenses and mirrors. The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.

@@

Figure \ref{fig:SingleElectron} shows the horizontal intensity profile for a filament beam (no emittances). It can be seen that there is little difference between using thin ideal lenses or the elliptical mirror model. The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in agreement with the findings of \cite{Canestrari2014}. Notice that the differences are minimal and to accentuate them we have chosen logarithmically scaled plots to represent the intensity distributions at the focal position. It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible - c.f. Fig.~5 in \cite{Canestrari2014}.  Increasing the complexity of the model increments considerably the simulation time, which went from typically $\sim16$~s to $\sim24$~s.

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_real.png}
    \label{fig:FiniteEmittanceA}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_real.png}
    \label{fig:FiniteEmittanceB}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice, using ideal focusing elements (top), and including the effect of the mirrors slope errors (bottom).
    }
\end{figure}

@@

This analysis of the mode transmission is more dramatic if one looks at the total intensity in the mode, which is given by its eigenvalue. Neglecting modes higher than 10, the lowest coherent mode ($m$=0) carries the 49.5\% of the beam intensity, 18.4\% for the mode $m$=2, and less than 5\% for the others.

@@

After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the X-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values obtained using more sophisticated methods. The size of the focused beam (e.g, 10$\times$5~nm$^2$ for EBS) and the transmission (3\%) calculated analytically are in good agreement with geometric ray-tracing (8.4$\times$4.8~nm$^2$, and 2.28\%). When characterizing beam sizes, one usually assumes that the intensity distributions are Gaussian and therefore they are well characterized by the FWHM value. From Fig.~\ref{fig:ray-tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF-1, as a result of cropping the beam with the VSS opened to 50$\mu$m. Although the electron beam statistics is in a very good approximation Gaussian, the undulator emission is not. In our ray-tracing simulations, we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source. This is certainly acceptable at the resonance but will fail out of resonance. Indeed, this lack of Gaussianity is relevant in the calculations of the coherent fraction. We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a filament electron beam in a magnetic field emits a fully coherent wavefront. However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam. This is justified considering that Eqs.~\ref{eq:photon small sigmas} are approximations obtained by fitting with Gaussians the theoretical non-Gaussian distributions (see Ref.~\cite{elleaume}). The correct definition of CF as the occupation of the lowest coherent mode does not present these incongruences. However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is very costly. It is, therefore, the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.

The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 12.56$\times$9.38~nm$^2$ SRW) and then add the slope error effect. The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 13.6$\times$9.14~nm$^2$ SRW). It is important to treat the slope errors correctly. Slope errors can be included in the ray-tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors. This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones. It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results. Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}. Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle). They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = k z(x) \sin \theta $. This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.

@@

\begin{figure}\label{fig:metrology}
\includegraphics[width=0.95\textwidth]{GRAPHICS/metrology.eps}
\caption{Error height profiles used for the simulations (left) and power spectral density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}
\end{figure}

>>>>

Accurate calculation and quantitative evaluation of the parameters related to X-ray optics are of great importance for designing, building and exploiting new beamlines. Every modern beamline follows a procedure of conceptual design plus technical design where detailed simulations are essential. These calculations allow testing of the design parameters in a virtual computer environment where advantages and limitations can be studied quantitatively. The selection of the optimal optical layout requires a detailed simulation of the optics imperfections, that in many cases are the limiting factor of the beamline optics. Deformations of the optical elements due to heat load need to be controlled, therefore optics simulations have to include results from engineering modeling of the thermal deformations usually done using finite element analysis. Traditionally the optics calculations for synchrotron beamlines are done using ray-tracing techniques. They model the X-ray beam as a collection of incoherent rays. This technique is still essential for obtaining information on aberrations, flux propagation, monochromator bandwidth, etc. However, the high coherence of the new sources can introduce the necessity to complement the ray-tracing methods with models including effects of diffraction and scattering from a coherent beam. A fully coherent hard X-ray beam is still a dream for the new generation storage rings. Despite a considerable increase in the coherent fraction for the upgraded sources, at hard X-ray photon energies it is still a few percent. Consequently, the ray-tracing analysis does not become obsolete but has to be complemented with an analysis based on wave optics. Because of the partial coherence, the wave optics analysis becomes nontrivial and can be done with different levels of approximation, as discussed in this work. In addition, wave optics simulations are usually computationally expensive.

@@

 In this paper, some quantitative results on how the storage ring upgrade will affect a particular beamline are presented. The paper is structured as follows: we first introduce in Section~\ref{Beamline description} the main parameters of the beamline and explain the particular choice of the parameters used in the calculations. Section~\ref{Simulations} is the central part of the paper, and describes the different methods used to analyze the beamline with increasing complexity, setting a hierarchy of methodologies that can be exported to the simulations of any beamline. In Section \ref{level0} we calculate with simple methods approximate values of focal size, flux and coherent fraction. Accuracy can be improved using ray-tracing methods in Section \ref{level1} that allow a more complete simulation of the beamline optics within certain limits which are related, in this example, to high demagnification and high coherence of the studied beamline. The wave optics section~\ref{level2} includes different models of increasing complexity as the required accuracy of the results becomes more demanding. In the Discussion (Section \ref{discussion}) the results of the different simulation methods are commented. The hierarchical approach allows more accurate and better results but the price to pay is to deal with more complex and computationally expensive tools. Finally, Section \ref{summary} summarizes the results.

@@

\begin{table}\label{tab:eBeam}
    \centering
    \caption{ESRF-1 and EBS main electron beam parameters. Electron beam parameters for the high-$\beta$ (labelled ESRF) and the EBS straight sections. Values taken at the symmetry point of the straight section, where the insertion devices are placed. Values are taken from \cite{orangebook}, and updated for EBS with
    the lattice labelled S28D. 
    %For EBS, two (very close) design values are used, from Ref.~\cite{ESRF2014} labelled EBS-S28A, and those coming from further lattice refinement, labelled EBS-S28D. 
    %\todo{Remove ESRF-AT and EBS-S28A} 
    }
        \resizebox{\textwidth}{!}{%
        \begin{tabular}{ccccc}
         Lattice & $\sigma_e$ horizontal {[}$\mu$m{]} &  $\sigma_e'$ horizontal {[}$\mu$rad{]} & $\sigma_e$ vertical {[}$\mu$m{]} & $\sigma_e'$ vertical {[}$\mu$rad{]} \\ \\
         %\input{tableElectron.txt}
         ESRF-1  & 387.80  & 3.50  & 10.30 & 1.20 \\ 
         %ESRF-AT  & 414.97  & 3.43  & 10.31 & 1.16 \\ 
         %EBS-S28A  & 27.20  & 3.40  & 5.20 & 1.40 \\ 
         EBS     & 30.18  & 3.64  & 4.37 & 1.37 \\ 
    \end{tabular}
    }
\end{table}

@@

The source size and divergence of the photon source are usually calculated approximately supposing Gaussian distribution of both the electron positions and divergences (Table \ref{tab:eBeam}) and the single-electron emission. In some cases, such as for undulator emission, this Gaussian approximation may not be accurate and have some implications that will be discussed later. Let $\sigma_{e(h,v)}$ and $\sigma_{e(h,v)}'$ be the standard deviations of electron beam size and divergence at the center of symmetry of the straight section in a plane perpendicular to the main propagation direction. The subscripts $h$ and $v$ stand for horizontal and vertical, respectively.

@@

Using the values for U18.3, $n$=1, and  $E$=17225~eV and the optical configuration described in Table \ref{tab:Positions}, one obtains a focal size 139$\times$5~nm$^2$ (H$\times$V) for ESRF-1 and $10\times5$~nm$^2$ for EBS (Table~\ref{tab:HandCalculations}). The horizontal focal spot for the ESRF-1 lattice is larger than the specified target value for the beamline, therefore the VSS could be closed to 50~$\mu$m thus reducing the beam size at the VSS plane by a factor ca. 8.
%\inred{8 $\approx$ 403.8/50=8.076}
Therefore, for ESRF-1, the horizontal size at the sample will be affected by the same factor obtaining roughly 17$\times$5~nm$^2$.

@@

To estimate the geometrical transmission of the beamline we have to consider which optical elements crop the beam and remove photons. This can be analyzed using their angular acceptance or numerical aperture. The multilayer monochromator accepts the whole beam. We consider a VSS closed to $S_h$=50~$\mu$m in horizontal and fully opened in vertical. For ESRF-1 its transmission is 1/8 (as discussed before) and 1 for EBS. In a similar way, the KB mirrors will crop the beam to a size that is the projection of their useful optical length onto a direction perpendicular to the optical axis. Using a mirror length of $L_v$=60~mm and $L_h$=26~mm for VFM and HFM, respectively, and incident angle of 15~mrad, the KB behaves as an aperture of  390 $\times$ 900 (H $\times$ V) $\mu$m$^2$. 
We get N.A. values of 2.7 $\times$ 4.9~$\mu$rad$^2$ (aperture size divided by the distance from the mirror to the VSS in H and to the source in V). The transmission coefficient is the ratio of the accepted N.A. over the beam divergence, resulting in 
4.1\% (H) and 40\% (V) for ESRF-1 and 7.1\% and 40\% for EBS. The overall transmission considers the product of H and V ratios, therefore 
0.21\% for ESRF-1 and 2.85\% for EBS, thus giving 
$4\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS. 
%\todo{check with references $10^9$ at 33.6 keV. I get: 1.6e+15 * 0.00295 * 0.7**3 = 1.62e+12. How closed are the slits? We used the undulator peak value over the whole ML bandwidth, that overestimates the integrated flux.}
%\todo{need to consider ML reflectivity for the 3 reflections, also undulator is not perfect, the horizontal slit may be closed to less than $50 \mu m$, also KB mirrors may not be fully illuminated and slope error losses due to initial ML mono are neglected}

@@

In summary, we have shown in this section how simple analytical calculations that can be done mostly ``by hand'' help to estimate the geometrical beam size at focal position (17$\times$5~nm$^2$ for ESRF-1, 10$\times$5~nm$^2$ for the EBS) the flux ($4.0\cdot10^{12}$ photons/s for ESRF-1 and $8.3\cdot10^{13}$ photons/s for EBS) and also anticipate that the beam is highly coherent producing a diffraction-limited focal spot of  7.75$\times$7.75~nm$^2$. Thus far we ignored the effect of real optics exhibiting surface errors, the reflectivity of the elements, and other aspects that will be treated in depth in the next sections. This is, however, the first step to be done when analyzing a beamline. It also helps in providing an initial benchmark estimate which should, to a large extent, provide a check of the validity of the numerical calculations that will provide more precise values, but should remain at the same order of magnitude.

@@

We performed ray-tracing of the ID16A beamline using an undulator source tuned to have its resonance at 17225~eV and approximated using Gaussian distributions (Eqs.~\ref{eq:photon big sigmas}), considering the ML as a cylindrical mirror, and using elliptical cylinder reflectors with finite dimensions as the KB focusing elements. The KB reflector shape are elliptical as required for a point-to-point focusing. For a first calculation, the reflectors (ML and KBs) are considered perfect and have no slope errors. 
The results in Fig.~\ref{fig:ray-tracing}a show two phenomena. One is that the migration from ESRF-1 lattice to the EBS will significantly improve the horizontal size, and the VSS is not needed. For EBS, the whole beam passes though the VSS, producing a final spot with almost Gaussian horizontal intensity profile instead of the step distribution for ESRF-1. Secondly, one can observe a much larger luminosity in the EBS case. The transmission values are 0.16\% for ESRF-1 and 2.22\% for EBS. These values confirm the analytical estimations (0.21\% for ESRF-1 and 2.85\% for EBS). As mentioned before, these values are similar to the coherent fraction \cite{arxivCF} (0.12 for ESRF-1 and 2.73\% for EBS), meaning that a very high coherence of the beam is expected. It is therefore expected some broadening of the beam due to diffraction effects originated by the mirrors acceptance.

The hybrid method \cite{hybrid} uses geometric ray-tracing combined with wavefront propagation. The code computes diffraction effects when the beam is cropped and can also simulate the effect of mirror figure errors when diffraction is present. The method can be applied to an entire beamline by simulating each optical element iteratively under the hypothesis of the validity of far-field propagation in all intermediate propagations. If this is not the case, a near field version is available for investigating imaging by individual optics. The interest of the hybrid method is to assess in a fast way the influence of beam coherence because the code is considerably faster than the full methods for dealing with the partial coherence of the synchrotron beam that are described below. Fig.~\ref{fig:ray-tracing}b presents the results using the hybrid model for the beamline with ideal elliptical optical surfaces. When compared with the pure ray-tracing in Fig.~\ref{fig:ray-tracing}a it can be appreciated that the spot is broadened. This is more evident in the vertical direction, because of the smaller focal size. The spot size is 8~nm in V for both storage ring lattices, and in horizontal is about 15~nm for ESRF-1 High$\beta$ and 11~nm for EBS. It is interesting to note the almost round spot that will be obtained with the new EBS lattice.

\begin{figure}
\label{fig:ray-tracing}
\centering
~~~~~ESRF-1~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~EBS
\newline
\includegraphics[width=0.95\textwidth]{GRAPHICS/figure3.png}
\caption{Image produced by different types of ideal focusing for the ESRF (left column) and EBS (right column). The source follows Gaussian distributions with parameters in Table~\ref{tab:HandCalculations}. 
a) Simulations using the ideal reflectors (cylindrical for ML, elliptical for the KB mirrors) without slope errors. 
b) Images calculated using the hybrid method (ray-tracing plus coherent diffraction) with ideal reflectors with no slope errors. It shows the focal spot degradation due to the diffraction arising from the cropping of the beam by KB mirrors. c) Simulations using the hybrid method including slope errors from the real mirror profiles measured at the ESRF metrology lab (see Appendix~\ref{appendix:metrology}). It can be observed further beam broadening and the growth of satellite structures. 
The focal beam dimensions (FWHM of the histograms) are written in the plot title.
}
\end{figure}

@@

A next step in the wave optics simulation is, assuming full coherence, to generate a 2D source wavefront that matches in detail the real source (an undulator in a storage ring) and then simulate the propagation of such a wavefront through the optical elements with models that reproduce well the characteristics of the real elements. Once the source is defined, several approaches of increasing complexity are used: a) use ideal thin lenses preceded by slits to emulate finite apertures (as done in the 1D model presented in Section~\ref{wofry}) b) use grazing incidence focusing optics models \cite{Canestrari2014}; and c) introduce the effect of imperfect optics on the simulations, i.e. slope and figure errors to both ideal lenses and mirrors. The computer code chosen for these calculations is ``Synchrotron Radiation Workshop'' (SRW) \cite{codeSRW}, which over the past two decades has been extensively benchmarked throughout different synchrotron radiation facilities and has become widely accepted within the X-ray community.

@@

Figure \ref{fig:SingleElectron} shows the horizontal intensity profile for a filament beam (no emittances). It can be seen that there is little difference between using thin ideal lenses or the elliptical mirror model. The elliptical mirror presents a slight reduction in intensity and asymmetry but no significant differences are observed, which is in agreement with the findings of \cite{Canestrari2014}. It is also important to mention that this equivalency between ideal lenses and elliptical mirrors is only valid close to the focal position: down- or upstream the focal position, differences are no longer negligible - c.f. Fig.~5 in \cite{Canestrari2014}.  Increasing the complexity of the model increments considerably the simulation time, which went from typically $\sim16$~s to $\sim24$~s.

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/ESRF_real.png}
    \label{fig:FiniteEmittanceA}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the ESRF-1 lattice. Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue). Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.
    }
\end{figure}

@@

\begin{figure}
    \centering
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_ideal.png}
        \includegraphics[width=0.95\textwidth]{GRAPHICS/EBS_real.png}
    \label{fig:FiniteEmittanceB}
    \caption{
    Intensity profile at the sample position produced by multi-electron method sampled with parameters of the EBS lattice. 
    Top: the perfect focusing with ideal lenses (green) is compared with the focusing using elliptical mirrors (blue). Bottom: the same as in the previous plot, but adding the effect of the mirrors slope errors in both cases.
    }
\end{figure}

@@

This analysis of the mode transmission is more dramatic if one looks at the total intensity in the mode, which is given by its eigenvalue. Neglecting modes higher than 10, the lowest coherent mode at the source ($m$=0) carries the 49.5\% of the beam intensity after being cropped by the beamline, 18.4\% for the mode $m$=2, and less than 5\% for the others.

@@

After the complete analysis done in the previous section on how the ID16A optics modify the characteristics of the X-ray beam, one can verify that the analytical previsions (Section~\ref{level0}) were a good estimator of the values obtained using more sophisticated methods. The size of the focused beam (e.g, 10$\times$5~nm$^2$ for EBS) and the transmission (2.85\%) calculated analytically are in good agreement with geometric ray-tracing (8.4$\times$4.8~nm$^2$, and 2.28\%). When characterizing beam sizes, one usually assumes that the intensity distributions are Gaussian and therefore they are well characterized by the FWHM value. From Fig.~\ref{fig:ray-tracing} one can see that the horizontal intensity distribution is not Gaussian for the ESRF-1, as a result of cropping the beam with the VSS opened to 50$\mu$m. Although the electron beam statistics is in a very good approximation Gaussian, the undulator emission is not. In our ray-tracing simulations, we always suppose that we work at the resonance, then we applied the Gaussian approximation used in Eqs.~\ref{eq:photon small sigmas} and \ref{eq:photon big sigmas} to compute the geometrical characteristics of the source. This is certainly acceptable at the resonance but will fail out of resonance. Indeed, this lack of Gaussianity is relevant in the calculations of the coherent fraction. We approximate the coherence fraction by Eq.~\ref{eq:coherent fraction}, which for zero emittance ($\sigma_{x,y}=\sigma_{x',y'}$=0) becomes CF=1, as expected because a filament electron beam in a magnetic field emits a fully coherent wavefront. However, the photon source for a filament beam would have an emittance of $\sigma_u \sigma_{u'} \approx \lambda / 2 \pi$ (using Eqs.~\ref{eq:photon small sigmas}), which is in contradiction with theory, that says that a pure coherent state has emittance of $\lambda / 4 \pi$, as happens for a Gaussian photon beam. This is justified considering that Eqs.~\ref{eq:photon small sigmas} are approximations obtained by fitting with Gaussians the theoretical non-Gaussian distributions (see Ref.~\cite{elleaume}). The correct definition of CF as the occupation of the lowest coherent mode does not present these incongruences. However, its calculation requires to perform the coherent mode decomposition, as done in Section~\ref{comsyl} which is very costly. It is, therefore, the case to find the compromise between accuracy and cost, or where in the hierarchical series of methodologies proposed one should stop.

The combination of ray-tracing and wave optics methods used indicate that the focalized beam size due to  demagnification with mirrors (giving for EBS 10$\times$5~nm$^2$ analytically, 8.8$\times$4.5~nm$^2$ with ray-tracing) has to be corrected with diffraction effect (11.2$\times$8~nm$^2$ hybrid, 11.6$\times$9.4~nm$^2$ SRW) and then add the slope error effect. The slope errors do not blow up the size and kit it in the specifications range (11.4$\times$8.5~nm$^2$ hybrid 12.5$\times$9.1~nm$^2$ SRW). It is important to treat the slope errors correctly. Slope errors can be included in the ray-tracing using the data in Appendix~\ref{appendix:metrology} and applying specular reflection on the optical surface with errors. This is the standard model of including slope errors in SHADOW, using measured profiles or synthetic ones. It works well for the case of incoherent beams, but in our case, because of the high coherence of the beam the ray-tracing method (specular reflection) overestimates the blowing of the focal spot conducting to wrong results. Therefore the slope errors of the KB system have been considered using a ``wave optics approach'', as explained in Ref.~\cite{hybrid}. Here, the mirror height errors $z(x)$ are projected to the plane perpendicular to the propagation ($z(x) \sin \theta$, with $\theta$ the grazing angle). They introduce a phase shift $\psi$ in the wavefront phase proportional to the optical path $\psi = k z(x) \sin \theta $. This is the algorithm used in the simulations in Sections~\ref{level1}, \ref{srw_se} and \ref{srw_me}.

@@

\begin{figure}\label{fig:metrology}
\includegraphics[width=0.9\textwidth]{GRAPHICS/metrology.eps}
\caption{Error height profiles used for the simulations (left) and power spectral density (right) for the ML monochromator (top row), the KB vertical focusing mirror (central row), and the KB horizontal focusing mirror (bottom row).}
\end{figure}
